### 2017-05-10 paper/paper.tex
More typos
Added description of metrics

@@

@@

@@

@@

The ballot data is first mapped to only include first preferences. Then the algorithm loops over the ballots and gets the sum of votes for each candidate. This process is shown in code example~\ref{lst:fptp processing}.
\codeblock{code/fptp/processing.ts}{lst:fptp processing}{Processing and formatting of data}
The results array is sorted and sliced to only include the $N$ candidates with most votes as seen in code example~\ref{lst:fptp result calculation}.
\codeblock{code/fptp/result.ts}{lst:fptp result calculation}{Finding and returning winning candidates}

@@

In order to eliminate and transfer surplus votes from any candidate node $A$, the children of $A$ are passed into the distribute function. In the case of transferring surplus votes, the votes for each child node is first multiplied by $\frac{\text{surplus votes}}{\text{total votes}}$ as outlined in the pseudocode in section~\ref{alg:stv psuedocode}. The full source code for functions that manipulate the tree structure can be found in \Code{/stv/tree.ts}.

The main process of the program is found in the \Code{stv/election.ts} file and runs the process defined in the pseudocode in section~\ref{alg:stv psuedocode} on page~\pageref{alg:stv psuedocode}.

\subsection{Schulze}
\subsubsection{Description}
The Schulze Method is a method used mainly to determine results in a single-winner election but can also be used to provide rankings and elect multiple candidates in a single election \autocite{Schulze2011}. Ballots are identical to those of the single transferable vote, see figure~\ref{fig:stv ballot} on page~\pageref{fig:stv ballot}. The method is compliant with the Condorcet Criterion, which means that if there is a candidate who the majority prefers in a pairwise comparison with every other candidate, that candidate wins. As mentioned, this method relies on comparing the preferences of every candidate to one another as shown in table~\ref{tab:pairwise comparison matrix}. If there is a Condorcet winner the process is simple: declare the Condorcet winner a winner and run another iteration without that candidate. Repeat this process until all seats are filled. However, there can be Condorcet ties, where there is no Condorcet winner. There are multiple ways of resolving the tie, one of them being the Schulze Method.

@@

The program then runs the algorithm outlined in the pseudocode in section~\ref{alg:schulze psuedocode}.

@@

\begin{table}
\centering
\caption{Key metrics for result in scenario 1}
\label{tab:scenario 1 result}
\begin{tabular}{@{}lccc@{}}
\toprule
Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
FPTP & 223 & 67\% & 2 \\
STV & 315 & 7\% & 2 \\
Schulze & 313 & 9\% & 2 \\ \bottomrule
\end{tabular}
\end{table}
% [ 223.39259164082, 315.44928276983, 313.6271671906]
\subsection{Scenario 2}
Scenario 2 featured 9 candidates of which 3 were to be elected. The distribution of votes was ``normal'', and as seen in figure~\ref{fig:example of scenario 2}, the election is very close with many candidates having a similar percentage of first-preference votes. The results obtained in scenario 2 are very interesting. FPTP tended to underrepresent candidate 1, and elected a candidate from ideology 2 only in 28\% of elections, despite the ideology theoretically having 30\% of popular support. The Schulze method also tended to underrepresent ideology 2. It instead overrepresented ideology 2, electing 3 candidates of ideology 2 in 50\% of elections. This indicates that the Schulze method is not a representative method in a multi-seat election. The STV result showed less misrepresentation than the FPTP result, which was to be expected since the method has access to more data of any voter's preferences. The results in scenario 2 can be observed in figure~\ref{fig:scenario 2 result} and the key metrics can be seen in table~\ref{tab:scenario 2 result}.
\begin{figure}
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width = 0.9\textwidth,
			height = 0.5\textwidth,
			legend style={at={(0.5,-0.25)},
			anchor=north,legend columns=-1},
      x tick style = transparent,
			ylabel = {Elections won},
			xlabel = {Candidate},
			ymajorgrids = true,
			every axis plot/.append style={draw=none,fill,no markers},
      ymin = 0,
			symbolic x coords={0,1,2,3,4,5,6,7,8},
			bar width=0.25cm
      ]
			\addplot coordinates
			{(0, 973)(1, 267)(2, 1000)(3, 0)(4, 12)(5, 743)(6, 0)(7, 5)(8, 0)};
			\addplot coordinates
			{(0, 456)(1, 808)(2, 1000)(3, 0)(4, 33)(5, 703)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 338)(1, 160)(2, 1000)(3, 0)(4, 0)(5, 1000)(6, 0)(7, 502)(8, 0)};
			\legend{FPTP, STV, Schulze}
		\end{axis}
	\end{tikzpicture}
	\caption{Results in scenario 2}
\label{fig:scenario 2 result}
\end{figure}

\begin{table}
\centering
\caption{Key metrics for result in scenario 2}
\label{tab:scenario 2 result}
\begin{tabular}{@{}lccc@{}}
\toprule
Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
FPTP & 418 & 42\% & 0, 2 \& 5 \\
STV & 388 & 14\% & 1, 2 \& 5 \\
Schulze & 393 & 67\% & 2, 5 \& 7  \\ \bottomrule
\end{tabular}
\end{table}
% [ 418.13628234977, 388.17206952015, 393.25535950293 ]
\subsection{Scenario 3}
Scenario 3 featured 9 candidates of which 3 were to be elected. The distribution of votes was steep. This can be seen by the first-preference votes in figure~\ref{fig:example of scenario 3}. The results obtained with each method in this scenario are practically identical. From this, it can be inferred that the results of the methods converge as the preferences in a population become more clear. The results obtained in this scenario can be seen in figure~\ref{fig:scenario 3 result}. The key metrics can be found in table~\ref{tab:scenario 3 result}.
\begin{figure}[H]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width = 0.9\textwidth,
			height = 0.5\textwidth,
			legend style={at={(0.5,-0.25)},
			anchor=north,legend columns=-1},
      x tick style = transparent,
			ylabel = {Elections won},
			xlabel = {Candidate},
			ymajorgrids = true,
			every axis plot/.append style={draw=none,fill,no markers},
      ymin = 0,
			symbolic x coords={0,1,2,3,4,5,6,7,8},
			bar width=0.25cm
      ]
			\addplot coordinates
			{(0, 994)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 6)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 1000)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 0)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 999)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 1)(6, 0)(7, 0)(8, 0)};
			\legend{FPTP, STV, Schulze}
			\end{axis}
	\end{tikzpicture}
	\caption{Results in scenario 3}
\label{fig:scenario 3 result}
\end{figure}

\begin{table}[H]
	\centering
	\caption{Key metrics for result in scenario 3}
\label{tab:scenario 3 result}
	\begin{tabular}{@{}lccc@{}}
		\toprule
		Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
		FPTP & 470 & 33\% & 0, 1 \& 2 \\
		STV & 471 & 33\% & 0, 1 \& 2 \\
		Schulze & 471 & 33\% & 0, 1 \& 2  \\ \bottomrule
	\end{tabular}
\end{table}
\section{Conclusion}
The results obtained in the three different scenarios provide some insight into the varying properties of the different voting systems. Each of the scenarios showed different behaviors for the different methods. For instance, the methods provided practically the same result in scenario 3 but wildly varying results in scenario 2. While the testing method used in this paper may not have been the most robust way of comparing the voting methods, several conclusions can be drawn from the data. I outline them in 4 points below.
\begin{enumerate}
	\item Schulze did not seem like a proportional voting method in multi-seat elections. ...
	\item In single-seat elections, Schulze and STV (and IRV in extension) seemed to provide similar results. ...
	\item FPTP did not seem like a very proportional method in close elections. ...
	\item STV seems to be the most proportional method out of the 3 tested, but is not without its flaws. ...
\end{enumerate}
Summary:
There is no perfect voting system. Generally, if you want proportional representation (which in a democracy you want), STV will likely be the best method to use out of the methods covered in this paper.
Modeling of data needs more work, as well as including other methods (such as \ldots). More research is needed.

@@

Hello

>>>>

\label{sec:method-scenario1}

@@

\label{sec:method-scenario2}

@@

\label{sec:method-scenario3}

@@

\section{Metrics for evaluating results}
Three metrics will be used for comparing the results obtained in the different scenarios. The results obtained with each voting method will be compared against the other methods. Comparisons to the configuration files outlined in Sections~\ref{sec:method-scenario1} -~\ref{sec:method-scenario3}.
\subsection{Standard deviation}
The standard deviation of the outcomes represents the variance in outcome due to the pseudorandom nature of the generator program. The standard deviation is calculated according to equation~\ref{eq:standard-deviation}. 
\begin{equation}
	\label{eq:standard-deviation}
	\sigma ={\sqrt {{\frac {1}{N}}\sum _{i=1}^{N}(x_{i}-\mu )^{2}}},{\rm {\ \ where\ \ }}\mu ={\frac {1}{N}}\sum _{i=1}^{N}x_{i}.
\end{equation}
\subsection{Misrepresentation}
The misrepresentation metric gives an idea of how \textit{proportional} the outcomes of a certain voting method in any scenario is. The level of proportionality of a result is a measure of how closely the preferences of a population are matched with the result. Higher proportionality means that the outcome more closely resembles the preferences of the voters. The method of calculating the misrepresentation used in this paper is defined as follows:
	\item Calculate what theoretical election outcome would be the most proportionate based on the configuration file used to generate the ballots.\label{enum:theoretical}
	\item Compare the actual outcomes of a certain method with the theoretical result obtained in Step \ref{enum:theoretical}.
		\item Calculate the difference between the measured and theoretical percentage for each ideology.
		\item Sum all the differences.
\subsection{Most likely winners}
Shows what candidates, given a certain scenario, are most likely to win in a particular method.

@@

The program sums all the first-preference votes for each candidate and elects the ones with most votes, in accordance with specified amount of seats.

@@

In order to eliminate and transfer surplus votes from any candidate node $A$, the children of $A$ are passed into the distribute function. In the case of transferring surplus votes, the votes for each child node is first multiplied by $\frac{\text{surplus votes}}{\text{total votes}}$ as outlined in the pseudocode in Section~\ref{alg:stv psuedocode}. The full source code for functions that manipulate the tree structure can be found in \Code{/stv/tree.ts}.
The main process of the program is found in the \Code{stv/election.ts} file and runs the process defined in the pseudocode in Section~\ref{alg:stv psuedocode} on page~\pageref{alg:stv psuedocode}.
\subsection{Schulze}
The Schulze Method is a method used mainly to determine results in a single-winner election but can also be used to provide rankings and elect multiple candidates in a single election \autocite{Schulze2011}. Ballots are identical to those of the single transferable vote, see figure~\ref{fig:stv ballot} on page~\pageref{fig:stv ballot}. The method is compliant with the Condorcet Criterion, which means that if there is a candidate who the majority prefers in a pairwise comparison with every other candidate, that candidate wins. As mentioned, this method relies on comparing the preferences of every candidate to one another as shown in table~\ref{tab:pairwise comparison matrix}. If there is a Condorcet winner the process is simple: declare the Condorcet winner a winner and run another iteration without that candidate. Repeat this process until all seats are filled. However, there can be Condorcet ties, where there is no Condorcet winner. There are multiple ways of resolving the tie, one of them being the Schulze Method.

@@

The program then runs the algorithm outlined in the pseudocode in Section~\ref{alg:schulze psuedocode}.

@@

\begin{table}
\centering
\caption{Key metrics for result in scenario 1}
\label{tab:scenario 1 result}
\begin{tabular}{@{}lccc@{}}
\toprule
Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
FPTP & 223 & 66 & 2 \\
STV & 315 & 8 & 2 \\
Schulze & 313 & 9 & 2 \\ \bottomrule
\end{tabular}
\end{table}
% [ 223.39259164082, 315.44928276983, 313.6271671906]
\subsection{Scenario 2}
Scenario 2 featured 9 candidates of which 3 were to be elected. The distribution of votes was ``normal'', and as seen in figure~\ref{fig:example of scenario 2}, the election is very close with many candidates having a similar percentage of first-preference votes. The results obtained in scenario 2 are very interesting. FPTP tended to underrepresent candidate 1, and elected a candidate from ideology 2 only in 28\% of elections, despite the ideology theoretically having 30\% of popular support. The Schulze method also tended to underrepresent ideology 2. It instead overrepresented ideology 2, electing 3 candidates of ideology 2 in 50\% of elections. This indicates that the Schulze method is not a representative method in a multi-seat election. The STV result showed less misrepresentation than the FPTP result, which was to be expected since the method has access to more data of any voter's preferences. The results in scenario 2 can be observed in figure~\ref{fig:scenario 2 result} and the key metrics can be seen in table~\ref{tab:scenario 2 result}.
\begin{figure}
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width = 0.9\textwidth,
			height = 0.5\textwidth,
			legend style={at={(0.5,-0.25)},
			anchor=north,legend columns=-1},
      x tick style = transparent,
			ylabel = {Elections won},
			xlabel = {Candidate},
			ymajorgrids = true,
			every axis plot/.append style={draw=none,fill,no markers},
      ymin = 0,
			symbolic x coords={0,1,2,3,4,5,6,7,8},
			bar width=0.25cm
      ]
			\addplot coordinates
			{(0, 973)(1, 267)(2, 1000)(3, 0)(4, 12)(5, 743)(6, 0)(7, 5)(8, 0)};
			\addplot coordinates
			{(0, 456)(1, 808)(2, 1000)(3, 0)(4, 33)(5, 703)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 338)(1, 160)(2, 1000)(3, 0)(4, 0)(5, 1000)(6, 0)(7, 502)(8, 0)};
			\legend{FPTP, STV, Schulze}
		\end{axis}
	\end{tikzpicture}
	\caption{Results in scenario 2}
\label{fig:scenario 2 result}
\end{figure}

\begin{table}
\centering
\caption{Key metrics for result in scenario 2}
\label{tab:scenario 2 result}
\begin{tabular}{@{}lccc@{}}
\toprule
Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
FPTP & 418 & 50 & 0, 2 \& 5 \\
STV & 388 & 47 & 1, 2 \& 5 \\
Schulze & 393 & 100 & 2, 5 \& 7  \\ \bottomrule
\end{tabular}
\end{table}
% [ 418.13628234977, 388.17206952015, 393.25535950293 ]
\subsection{Scenario 3}
Scenario 3 featured 9 candidates of which 3 were to be elected. The distribution of votes was steep. This can be seen by the first-preference votes in figure~\ref{fig:example of scenario 3}. The results obtained with each method in this scenario are practically identical. From this, it can be inferred that the results of the methods converge as the preferences in a population become more clear. The results obtained in this scenario can be seen in figure~\ref{fig:scenario 3 result}. The key metrics can be found in table~\ref{tab:scenario 3 result}.
\begin{figure}[H]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width = 0.9\textwidth,
			height = 0.5\textwidth,
			legend style={at={(0.5,-0.25)},
			anchor=north,legend columns=-1},
      x tick style = transparent,
			ylabel = {Elections won},
			xlabel = {Candidate},
			ymajorgrids = true,
			every axis plot/.append style={draw=none,fill,no markers},
      ymin = 0,
			symbolic x coords={0,1,2,3,4,5,6,7,8},
			bar width=0.25cm
      ]
			\addplot coordinates
			{(0, 994)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 6)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 1000)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 0)(6, 0)(7, 0)(8, 0)};
			\addplot coordinates
			{(0, 999)(1, 1000)(2, 1000)(3, 0)(4, 0)(5, 1)(6, 0)(7, 0)(8, 0)};
			\legend{FPTP, STV, Schulze}
			\end{axis}
	\end{tikzpicture}
	\caption{Results in scenario 3}
\label{fig:scenario 3 result}
\end{figure}

\begin{table}[H]
	\centering
	\caption{Key metrics for result in scenario 3}
\label{tab:scenario 3 result}
	\begin{tabular}{@{}lccc@{}}
		\toprule
		Method & Standard deviation & Misrepresentation & Most likely winners \\ \midrule
		FPTP & 470 & 0 & 0, 1 \& 2 \\
		STV & 471 & 0 & 0, 1 \& 2 \\
		Schulze & 471 & 0 & 0, 1 \& 2  \\ \bottomrule
	\end{tabular}
\end{table}
\section{Conclusion}
The results obtained in the three different scenarios provide some insight into the varying properties of the different voting systems. Each of the scenarios showed different behaviors for the different methods. For instance, the methods provided practically the same result in scenario 3 but wildly varying results in scenario 2. While the testing method used in this paper may not have been the most robust way of comparing the voting methods, several conclusions can be drawn from the data. I outline them in 4 points below.
\begin{enumerate}
	\item Schulze did not seem like a proportional voting method in multi-seat elections. ...
	\item In single-seat elections, Schulze and STV (and IRV in extension) seemed to provide similar results. ...
	\item FPTP did not seem like a very proportional method in close elections. ...
	\item STV seems to be the most proportional method out of the 3 tested, but is not without its flaws. ...
\end{enumerate}
Summary:
There is no perfect voting system. Generally, if you want proportional representation (which in a democracy you want), STV will likely be the best method to use out of the methods covered in this paper.
Modeling of data needs more work, as well as including other methods (such as \ldots). More research is needed.

@@

Hello world
